---
title: "Introduction to AlphaGenomeR"
author: "Himanshu Bhandary"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Introduction to AlphaGenomeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignettePackage{AlphaGenomeR}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE # Set to FALSE for Bioconductor builds (no internet access)
)
```

## Introduction

`AlphaGenomeR` is an R interface to Google DeepMind's AlphaGenome API. AlphaGenome is a multimodal transformer model that predicts functional genomic features (such as gene expression, chromatin accessibility, and splicing) from DNA sequences at single-base resolution.

This package bridges the official Python SDK using `reticulate` to enable high-throughput gRPC-based predictions directly within R.

## Prerequisites

Before using `AlphaGenomeR`, you must have:

1. **Python (>= 3.10)** installed.
2. The **AlphaGenome Python package**:
   ```bash
   pip install alphagenome
   ```
3. An **API Key** from the [AlphaGenome Science Page](https://deepmind.google/science/alphagenome/).

## Basic Usage

### Loading the Package

```{r}
library(AlphaGenomeR)
```

### Querying Predictions

To query predictions, provide your API key and a 1MB genomic region (hg38). Note that `ontology_terms` (UBERON or CL terms) are required for tissue-specific predictions.

```{r}
api_key <- "YOUR_API_KEY"
region <- "chr17:42560601-43609177"

# Query for Lung tissue
results <- alphagenome_query(
  access_token = api_key,
  genomic_region = region,
  ontology_terms = c("UBERON:0002048"),
  requested_outputs = c("RNA_SEQ", "ATAC")
)
```

## Extracting Data

`AlphaGenomeR` provides specialized extractor functions to convert the raw API response into R-native matrices and data frames.

### RNA-seq Data

```{r}
rna_data <- alphagenome_get_rna_seq(results)

# The extractor returns a list with 'values' and 'metadata'
dim(rna_data$values) # Matrix of predictions (Positions x Tracks)
head(rna_data$metadata) # Metadata for each track
```

### ATAC-seq Data

```{r}
atac_data <- alphagenome_get_atac(results)
# Access signal values
plot(atac_data$values[,1], type = "l", main = "Predicted ATAC-seq Signal")
```

## Supported Modalities

The following extractor functions are available:

* `alphagenome_get_rna_seq()`
* `alphagenome_get_atac()`
* `alphagenome_get_cage()`
* `alphagenome_get_dnase()`
* `alphagenome_get_chip_tf()`
* `alphagenome_get_chip_histone()`
* `alphagenome_get_splice_sites()`
* `alphagenome_get_splice_junctions()`
* `alphagenome_get_splice_usage()`
* `alphagenome_get_procap()`
* `alphagenome_get_contact_maps()`

## Session Info

```{r, eval=TRUE}
sessionInfo()
```
